---
output: reprex::reprex_document
knit: reprex::reprex_render
---


# import libraries
```{r}
library(dplyr)
```
# Import data
```{r}
rai <- read.csv("../data/rai_with_covariates.csv")
noc <- read.csv("../data/noct_with_covariates.csv")
project_cov <- read.csv("../data/project_covariates.csv")
combine <- read.csv("../data/COMBINE_data/trait_data_imputed.csv")
```

# Variables
Dietary breadth: Dietary breadth can be used as an indicator of the number of different food elements a species consumes. Estimated as the number of different EltonTraits categories that constitute ≥20% of a species’ diet (Usui et al. 2017).

Habitat breadth: Habitat breadth can be used as an indicator of a species’ environmental tolerance. Estimated as the number of distinct level 1 IUCN habitats suitable for the species.




```{r}
# reduce the data of combine to match the porjects data
combine_filtered <- combine %>%
  semi_join(rai, by = c("iucn2020_binomial" = "Species"))

# how many species in project?
n_distinct(rai$Species) #163

#found: 160
nrow(combine_filtered)

#which are not included?
species_not_in_combine <- rai %>%
  anti_join(combine, by = c("Species" = "iucn2020_binomial"))

unique_species_not_in_combine <- unique(species_not_in_combine$Species)
print(unique_species_not_in_combine)

```

# Calculation of relative brain size
To calculate relative brain size we divided log-transformed brain mass by log-transformed body mass
```{r}
species_df <- data.frame(
  species_iucn = combine$iucn2020_binomial,
  species_phylacine = combine$phylacine_binomial, 
  adult_mass_g = combine$adult_mass_g,
  brain_mass_g = combine$brain_mass_g
)

species_df$log_mass <- log(species_df$adult_mass_g)
species_df$log_brain <- log(species_df$brain_mass_g)
species_df$rel_br_size <- (species_df$log_brain / rel_br_size_df$log_mass)

# reduce data to match the data in project but three
species_df <- species_df %>%
  semi_join(rai, by = c("species_iucn" = "Species"))

# Ensure there are no duplicate species in rai
rai_unique <- rai %>%
  distinct(Species, .keep_all = TRUE)

# Add column with original data
species_df <- species_df %>%
  left_join(rai_unique %>% select(Species, brain_mass_relative), by = c("species_iucn" = "Species"))

#THIS WORKS

# create three equal sized groups by mass
species_df <- species_df %>%
  mutate(mass_group = ntile(adult_mass_g, 3))

break_values <- quantile(species_df$mass_kg, probs = seq(0, 1, length.out = 4))
break_values

median(species_df$adult_mass_g) #is roughly 11 kg

# plot the distribution of mass of projects
species_df$mass_kg <- species_df$adult_mass_g / 1000
hist(
  species_df$mass_kg, 
  breaks = seq(0, max(species_df$mass_kg, by = 20)),
  probability = FALSE) # to ensure absolute values


```

# Whats the variance in rai? - lets check on one example
```{r}
rai_test <- rai %>%
  filter(Species == "Alces alces")

median(rai_test$yi_mod_lowhigh)

```


```{r}

library(dplyr)
library(ggplot2)
data_paper <- read.csv("data/rai_with_covariates.csv", sep = ",")
data_combine <- read.csv("data/COMBINE_data/trait_data_imputed.csv")



data_paper <- data_paper %>%
  left_join(data_combine %>% select(iucn2020_binomial, adult_mass_g), by = c("Species" = "iucn2020_binomial"))

data_paper$weight_kg <- as.integer(data_paper$adult_mass_g / 1000)

# Remove outliers from adult_mass_g
Q1 <- quantile(data_paper$weight_kg, 0.25, na.rm = TRUE)
Q3 <- quantile(data_paper$weight_kg, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

data_combine_filtered <- data_paper %>%
  filter(weight_kg >= lower_bound & weight_kg <= upper_bound)
  
```
```{r}
# Plot the distribution of adult_mass_g without outliers
ggplot(data_paper, aes(x = weight_kg)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "black") 
```